Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-11 paper_enahpc14/Sec2:COSMO-ART.tex
restructures paper; improves contents

Thus, we give here a brief  overview of the main features of the model
setup for  our benchmark's daily  runs.  Three-dimensional simulations
are performed over  Europe for April 13th 2010, which  is close to the
equinox and thus  brings benefits of having approximately  half day of
night  and half  day  of  sun exposure,  therefore  ensuring a  proper
activation of the chemistry cycle. We consider daily 24-hour forecasts
without  any  previous spin-up  simulations.   The calculation  domain
corresponds to  the CORDEX-EU-44  domain and is  covered by a  grid of
$222\times   216$    points   with   a    horizontal   resolution   of
$0.22\,^{\circ}$,  i.e.  50  km in  both directions,  and  40 vertical
layers.   Boundary  and initial  conditions,  emmissions and  external
parameters  were obtained by  the INT2LM  pre-processing interpolation
tool, which interpolate  data from a coarse-grid driving  model to the
rotated    latitude-longitude   grid    of   the    COSMO-Model,   and
INT2LM\_trcr\_ART,  which  interpolates  aerosol  or  gaseous  species
initial and boundary conditions  and consists mostly of the definition
of a large number of  tracers.  The meteorological initial and bounday
conditions are  obtained by  the the ECMWF  global spectral  model IFS
with an update  frequency of 3h.  Boundary data  for gas-phase species
are taken from IFS-MOZART output at 6h temporal resolution.  The model
setup incorporates  34 2-d and 45  3-d fields to be  written out every
hour and doesn't include data assimilation methods.

>>>>

Thus, we give here a brief  overview of the main features of the model
setup for  our benchmark's daily  runs.  Three-dimensional simulations
are performed over  Europe for April 13th 2010, which  is close to the
equinox and thus  brings benefits of having approximately  half day of
night  and half  day  of  sun exposure,  therefore  ensuring a  proper
activation of the chemistry cycle. We consider daily 24-hour forecasts
without  any  previous spin-up  simulations.   The calculation  domain
corresponds to  the CORDEX-EU-44  domain and is  covered by a  grid of
$222\times   216$    points   with   a    horizontal   resolution   of
$0.22\,^{\circ}$,  i.e.  50  km in  both directions,  and  40 vertical
layers.   Boundary  and  initial  conditions, emissions  and  external
parameters  were obtained by  the INT2LM  pre-processing interpolation
tool, which interpolate  data from a coarse-grid driving  model to the
rotated    latitude-longitude   grid    of   the    COSMO-Model,   and
INT2LM\_trcr\_ART,  which  interpolates  aerosol  or  gaseous  species
initial and boundary conditions  and consists mostly of the definition
of a large number of  tracers. In particular when using COSMO-ART, the
following input data are necessary:

\begin{itemize}
\item  Gas phase:  Anthropogenic emissions  for different  species and
  land use data for biogenic emissions and deposition,
\item Aerosol particles: Anthropogenic emissions,
\item Mineral dust: Soil specific land use data.
\end{itemize}

The meteorological initial and  bounday conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at   6h  temporal   resolution.   NAMELIST-input   specifying  runtime
parameters    splitted   into    several   groups    are    given   in
Table~\ref{tab:1}.  The model  setup incorporates  34 2-d  and  45 3-d
fields  to  be  written  out  every  hour  and  doesn't  include  data
assimilation methods.

\begin{table}[htbf]
  \begin{center}
    \caption{}
    \label{tab:1}
    \begin{tabular}{ll}
      \hline\noalign{\smallskip} 
      \textbf{Group} & \textbf{Description} \\
      \noalign{\smallskip}\hline\noalign{\smallskip}
      LMGRID & Grid domain and size parameters \\
      RUNCTL & Model run parameters \\
      TUNING & Physics and dynamics parameters \\
      DYNCTL & Adiabatic model parameters \\
      PHYCTL & Diabatic model parameters \\
      COSMO\_ART & Gases and aerosols model parameters \\
      DIACTL & Diagnostic calculations parameters \\
      SATCTL & Synthetic satellite images parameters \\
      IOCTL & I/O environment parameters \\
      GRIBIN & GRIB input parameters \\
      GRIBOUT & GRIB output parameters \\
     \noalign{\smallskip}\hline
    \end{tabular}
  \end{center}
\end{table}

### 2014-04-11 paper_enahpc14/Sec3:Measurement-framework.tex
restructures paper; improves contents

\subsection{Experimental Platforms}
\label{subsec:2.1}
While hardware  platforms mature  and are replaced,  we have  chosen a
state-of-the-art  Intel's third-generation  Core (aka  ``Ivy Bridge'')
processing platform (called ``Monch'')  for our power measurements, as
it  is slated  to stay  in service  without hardware  upgrade  for the
duration  of  the Exa2Green  project.   In  principle  at least,  this
architecture could be recreated or found in an identical configuration
beyond the lifetime of the project.  Given that the baseline benchmark
can    be    reproduced    within    an   expected    variance    (see
subsection~\ref{subsec:1.2}), and that  the baseline run configuration
can be used in all future versions of the code, a fair comparison will
be made between the baseline  and the milestone versions of COSMO-ART.
In  this   study,  a  complementary   energy-to-solution  benchmarking
comparison is carried out on  the ``Pilatus'' cluster based on Intel's
previous-generation  ``Sandy  Bridge'' processors,  known  to be  more
power consuming.

\subsubsection{Monch (CSCS - ETH Zurich)}
Monch, which is installed  at the Swiss National Supercomputing Center
(CSCS)  of the ETH  Zurich, is  a 10  rack NEC-provided  (LX-2400) and
dual-socket Intel-based cluster, utilised from people that are part of
the   Swiss  Platform   for  Advanced   Scientific   Computing  (PASC,
\url{http://www.pasc-ch.org/}).   It  is   composed  of  312  standard
compute  nodes,  24  large-memory  compute nodes  and  24  huge-memory
compute nodes.   Each standard compute node comprises  2 Intel 10-core
Ivy Bridge  EP E5-2660v2 processors  operating at 2.2  GHz, themselves
connected by a high speed Infiniband network, based on Mellanox SX6036
managed  FDR switches,  with a  56  Gb/s speed.  Each core  has 32  KB
instruction and 32 KB  data L1 caches and 256 KB of  L2 cache. All the
10 core  share a 25.6 MB  L3 cache and  the platform has 32GB  of DDR3
1600MHz RAM.

\subsubsection{Pilatus (CSCS - ETH Zurich)}
Pilatus is a dual-socket Intel  Sandy-Bridge based cluster used as Piz
Daint pre-post processing cluster.  It is composed of 42 compute nodes
with  a total  number of  1344 cores  in hyper-threading  and  2688 GB
aggregate memory and has 2  high-speed interconnects based on FDR: the
first is  dedicated to the MPI  traffic and the second  to the storage
high  speed traffic.   The 2  login nodes  and the  42  computes nodes
consists in  11 twin-pair Intel  E5-Series DALCO r2264i4t  2U scalable
compute  modules.  Each  module contains  4 compute  nodes based  on 2
Intel  8-core Intel  Xeon  E5-2670 processors  operating  at 2.6  GHz,
themselves  connected by  a high  speed Infiniband  network,  based on
Mellanox SX6036 managed FDR switches,  with a 56 Gb/s speed. Each core
has  32 KB  instruction and  32 KB  data L1  caches and  256 KB  of L2
cache. All the 8 core share a 20 MB L3 cache and the platform has 64GB
of DDR3 1600MHz RAM.

\subsection{Power-Performance Measurement}
\label{subsec:2.2}

\subsubsection{CSCS - ETH Zurich}
Hardware platforms in our experiment include E3METER Intelligent Power
Strips  (IPS) and Monitors  (IPM) which  are high  quality electricity
meters  that enable  to  monitor and  optimize  energy consumption  of
datacenters or large facilities.  A  central data logger - the E3METER
Data  Concentrator  - aggregates  all  measured  data  and makes  them
available in  various forms (SNMP,  Web, etc). Data  transfers between
E3METER IPS and E3METER Concentrator use reliable narrowband powerline
communication  (PLC)  technology  which  avoids  the  need  for  extra
cabling.   It  has one  internal  temperature  sensor.  Two  dedicated
extension  ports  can be  used  to  measure  external temperature  and
humidity  through E3METER remote  sensors.  It  features non-intrusive
voltage, current and frequency measurement through current transformer
with an accuracy of 1\%.

>>>>

In this section,  we present two different frameworks  deployed on HPC
systems  to  measure the  power  consumption  and  performance of  the
baseline execution.

\subsection{Framework description}
\label{subsec:3.1}

\subsubsection{E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center of  ETH Zurich (CSCS) are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high  quality  electricity meters  released  by  Riedo Networks,  that
enable to monitor  and log power consumption of  the IT infrastructure
as  well as  constantly analyze  line voltage,  current, power-factor,
frequency  with  1\% accuracy.   Using  reliable narrowband  powerline
communication (PLC)  technology, all  metering and power  quality data
from   each  IPS  are   centrally  collected   by  the   E3METER  Data
Concentrator, via the existing power cables thus avoiding the need for
extra cabling.   This data  is made available  via SNMP,  HTTP, TELNET
through  the built-in  Fast  Ethernet port.   Time synchronization  is
guaranteed by  using NTP servers.   Measured data is  accessed through
the open source  Cacti software including the E3METER  Cacti Plugin to
scan the entire PLC network  and monitor the power usage of individual
rack in real-time.
